---
title: "Summary Statistics of PIRGd in WY using DLC"
author: "Ethan Berman"
date: "3/3/2020"
output: html_document
---

```{r, setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '/Volumes/SSD/climate_effects')

```

## Overview

This document contains summary statistics of PIRGd in WY from 2002-2018 from the DLC phenology dataset. Stats are calculated over the entire state and a smaller sub-region in SW Wyoming corresponding to sagebrush treatments.

## Mean PIRGd in Wyoming and SW Wyoming

```{r, echo=FALSE, message=FALSE, warning = F}
#load packages
library(rgdal)
library(sp)
library(raster)
library(tidyverse)
library(rasterVis)
library(viridis)

#load wyoming shapefile
wy <- readOGR('reference/wyoming.shp', verbose = FALSE)

#transformto CRS of emodis
wy <- spTransform(wy, "+proj=laea +lat_0=45 +lon_0=-100 +x_0=0 +y_0=0 +a=6370997 +b=6370997 +units=m +no_defs")

#load PIRGd brick
pirgd <- brick('dlc/maxIRGdate_wy_laea_2000_2019.tif')

#only use 2002 to 2018 to be consistent with eMODIS
pirgd <- pirgd[[3:19]]

#mask to WY before calc summary stats
pirgd <- mask(pirgd, wy)

#calc mean pirgd for each pixel
pirgd_m <- calc(pirgd, function(x) {mean(x,na.rm = T)})

#calc number of na values in annual layers
pirgd_na <- calc(pirgd, function(x) {sum(is.na(x))})

#only save pirgd mean value if less than half of years are missing
#we have 17 years total
pirgd_m[pirgd_na > 8] <- NA

#round to whole DOY
pirgd_m <- round(pirgd_m)

#write pirgd mean raster
#writeRaster(pirgd_m, './emodis/pirgd/pirgd_mean_wy_2002_2018.tif', format = "GTiff", overwrite = T)

#load shapefile of treatments in SW WY
treat <- readOGR('/Users/Ediz/Team Braintree Dropbox/Ethan Berman/USGS/forage project/Phase 2/Initial Treatment Polygons/Project_Plannig_Areas_SWRMDCopy2.shp', verbose = FALSE)

#transform to CRS of emodis
treat <- spTransform(treat, "+proj=laea +lat_0=45 +lon_0=-100 +x_0=0 +y_0=0 +a=6370997 +b=6370997 +units=m +no_defs")

#plot treatments in WY
plot(wy, main = "Treatment Polygons in SW Wyoming")
plot(treat, add = T)

#calculate treatment bounding box
treat_bb <- bbox(treat)
treat_bb <- as(raster::extent(treat_bb), "SpatialPolygons")
proj4string(treat_bb) <- "+proj=laea +lat_0=45 +lon_0=-100 +x_0=0 +y_0=0 +a=6370997 +b=6370997 +units=m +no_defs"

#buffer at 2 km
treat_bb <- buffer(treat_bb, width = 2000)

#crop to WY
treat_bb <- crop(treat_bb, wy)

#plot mean PIRGd in WY
## Set up color breaks
my_at <- seq(0, 250, 10)
my_ticks <- seq(0, 250, 50)

my_col <- list(at=my_at, ## where the colors change
                   labels=list(
                     at=my_ticks ## where to print labels
                     ))
col <- viridis(length(my_at)-1, direction = -1)

#plot
levelplot(pirgd_m, margin = F, col.regions = col, colorkey = my_col,
          main = list('Mean PIRGd in Wyoming from 2002-2018'),
          ylab.right = "DOY",
          par.settings = list(layout.widths = list(axis.key.padding = 0,
                                                     ylab.right = 2))) +
        layer(sp.lines(wy, col = 'black')) +
        layer(sp.lines(treat_bb, col = 'black'))

#plot histogram of mean pirgd
hist(pirgd_m[pirgd_m >= 0 & pirgd_m <= 250], main = "Mean PIRGd in Wyoming from 2002-2018", xlab = "DOY",
     xlim = c(0, 250), breaks = seq(0,250,10))

#plot boxplot of mean pirgd
boxplot(pirgd_m, ylab = "DOY", main = "Boxplot of Mean PIRGd dates in WY from 2002-2018",
        ylim = c(0, 250))

#crop mean pirgd to this bb
pirgd_m_bb <- crop(pirgd_m, treat_bb)

#plot pirgd mean raster for SW Wyoming
levelplot(pirgd_m_bb, margin = F, col.regions = col, colorkey = my_col,
          main = list('Mean PIRGd in SW Wyoming from 2002-2018'),
          ylab.right = "DOY",
          par.settings = list(layout.widths = list(axis.key.padding = 0,
                                                     ylab.right = 2)))

#plot histogram of mean pirgd
hist(pirgd_m_bb[pirgd_m_bb >= 0 & pirgd_m_bb <= 250], main = "Mean PIRGd in SW Wyoming from 2002-2018", xlab = "DOY",
     xlim = c(0, 250), breaks = seq(0,250,10))

#remove variables
rm(pirgd_m, pirgd_m_bb)

```

## Median PIRGd in Wyoming and SW Wyoming

```{r, echo=FALSE, message=FALSE}
#calc median pirgd for each pixel
pirgd_med <- calc(pirgd, function(x) {median(x,na.rm = T)})

#only save pirgd mean value if less than half of years are missing
#we have 17 years total
pirgd_med[pirgd_na > 8] <- NA

#round to whole DOY
pirgd_med <- round(pirgd_med)

#write pirgd median raster
#writeRaster(pirgd_med, './emodis/pirgd/pirgd_median_wy_2002_2018.tif', format = "GTiff", overwrite = F)

#plot median PIRGd in WY
levelplot(pirgd_med, margin = F, col.regions = col, colorkey = my_col,
          main = list('Median PIRGd in Wyoming from 2002-2018'),
          ylab.right = "DOY",
          par.settings = list(layout.widths = list(axis.key.padding = 0,
                                                     ylab.right = 2))) +
        layer(sp.lines(wy, col = 'black')) +
        layer(sp.lines(treat_bb, col = 'black'))

#plot histogram of median pirgd
hist(pirgd_med[pirgd_med >= 0 & pirgd_med <= 250], main = "Median PIRGd in Wyoming from 2002-2018", xlab = "DOY",
     xlim = c(0, 250), breaks = seq(0,250,10))

#plot boxplot of median pirgd
boxplot(pirgd_med, ylab = "DOY", main = "Boxplot of Median PIRGd dates in WY from 2002-2018",
        ylim = c(0, 250))

#crop median pirgd to this bb
pirgd_med_bb <- crop(pirgd_med, treat_bb)

#plot pirgd median raster for SW Wyoming
levelplot(pirgd_med_bb, margin = F, col.regions = col, colorkey = my_col,
          main = list('Median PIRGd in SW Wyoming from 2002-2018'),
          ylab.right = "DOY",
          par.settings = list(layout.widths = list(axis.key.padding = 0,
                                                     ylab.right = 2)))

#plot histogram of median pirgd
hist(pirgd_med_bb[pirgd_med_bb >= 0 & pirgd_med_bb <= 250], main = "Median PIRGd in SW Wyoming from 2002-2018", xlab = "DOY",
     xlim = c(0, 250), breaks = seq(0,250,10))

#remove variables
rm(pirgd_med, pirgd_med_bb)

```

## Variance of PIRGd in Wyoming and SW Wyoming

Variance > 5000 set to NA.

```{r, echo=FALSE, message=FALSE}
#calc var pirgd for each pixel
pirgd_var <- calc(pirgd, function(x) {var(x,na.rm = T)})

#only save pirgd var value if less than half of years are missing
#we have 17 years total
pirgd_var[pirgd_na > 8] <- NA

#round to whole DOY
pirgd_var <- round(pirgd_var)

pirgd_var2 <- pirgd_var
pirgd_var2[pirgd_var2 > 5000] <- NA
#write pirgd var raster
#writeRaster(pirgd_var, './emodis/pirgd/pirgd_var_wy_2002_2018.tif', format = "GTiff", overwrite = F)

#plot var PIRGd in WY
## Set up color breaks
my_at <- seq(0, 4000, 250)
my_ticks <- seq(0, 4000, 500)

my_col <- list(at=my_at, ## where the colors change
                   labels=list(
                     at=my_ticks ## where to print labels
                     ))
col <- viridis(length(my_at)-1, direction = -1)

#plot
levelplot(pirgd_var2, margin = F, col.regions = col, colorkey = my_col,
          main = list('Variance of PIRGd in Wyoming from 2002-2018'),
          ylab.right = "DOY",
          par.settings = list(layout.widths = list(axis.key.padding = 0,
                                                     ylab.right = 2))) +
        layer(sp.lines(wy, col = 'black')) +
        layer(sp.lines(treat_bb, col = 'black'))

#plot histogram of var pirgd
hist(pirgd_var2, main = "Variance of PIRGd in Wyoming from 2002-2018", xlab = "Days",
     xlim = c(0, 5000), breaks = seq(0,5000,100))

#crop var pirgd to this bb
pirgd_var_bb <- crop(pirgd_var2, treat_bb)

#plot pirgd var raster for SW Wyoming
levelplot(pirgd_var_bb, margin = F, col.regions = col, colorkey = my_col,
          main = list('Variance of PIRGd in SW Wyoming from 2002-2018'),
          ylab.right = "DOY",
          par.settings = list(layout.widths = list(axis.key.padding = 0,
                                                     ylab.right = 2))) +
        layer(sp.lines(wy, col = 'black')) +
        layer(sp.lines(treat_bb, col = 'black'))

#plot histogram of var pirgd
hist(pirgd_var_bb, main = "Variance of PIRGd in SW Wyoming from 2002-2018", xlab = "Days",
     xlim = c(0, 5000), breaks = seq(0,5000,100))

#remove variables
rm(pirgd_var, pirgd_var2, pirgd_var_bb)
```

## Mid-Range of PIRGd in Wyoming and SW Wyoming

Min + (Max-Min)/2

```{r, echo=FALSE, message=FALSE}
#calc mid pirgd for each pixel
pirgd_max <- calc(pirgd, function(x) {
        if(sum(is.na(x)) < 17) {max(x, na.rm = T)} else{NA}
        })

pirgd_min <- calc(pirgd, function(x) {
        if(sum(is.na(x)) < 17) {min(x, na.rm = T)} else{NA}
        })

pirgd_midrange <- pirgd_min + (pirgd_max - pirgd_min)/2

#only save pirgd midrange value if less than half of years are missing
#we have 17 years total
pirgd_midrange[pirgd_na > 8] <- NA

#round to whole DOY
pirgd_midrange <- round(pirgd_midrange)

#write pirgd midrange raster
#writeRaster(pirgd_midrange, '/Volumes/SSD/climate_effects/emodis/pirgd/pirgd_midrange_wy_2002_2018.tif', format = "GTiff", overwrite = F)

#plot midrange PIRGd in WY
## Set up color breaks
my_at <- seq(0, 250, 10)
my_ticks <- seq(0, 250, 50)

my_col <- list(at=my_at, ## where the colors change
                   labels=list(
                     at=my_ticks ## where to print labels
                     ))
col <- viridis(length(my_at)-1, direction = -1)

levelplot(pirgd_midrange, margin = F, col.regions = col, colorkey = my_col,
          main = list('Midrange of PIRGd in Wyoming from 2002-2018'),
          ylab.right = "DOY",
          par.settings = list(layout.widths = list(axis.key.padding = 0,
                                                     ylab.right = 2))) +
        layer(sp.lines(wy, col = 'black')) +
        layer(sp.lines(treat_bb, col = 'black'))

#plot histogram of midrange pirgd
hist(pirgd_midrange[pirgd_midrange >= 0 & pirgd_midrange <= 250], main = "Midrange of PIRGd in Wyoming from 2002-2018", xlab = "DOY",
     xlim = c(0, 250), breaks = seq(0,250,10))

#plot boxplot of midrange pirgd
boxplot(pirgd_midrange, ylab = "DOY", main = "Boxplot of Midrange of PIRGd dates in WY from 2002-2018",
        ylim = c(0, 250))

#crop midrange pirgd to this bb
pirgd_midrange_bb <- crop(pirgd_midrange, treat_bb)

#plot pirgd midrange raster for SW Wyoming
levelplot(pirgd_midrange_bb, margin = F, col.regions = col, colorkey = my_col,
          main = list('Midrange of PIRGd in SW Wyoming from 2002-2018'),
          ylab.right = "DOY",
          par.settings = list(layout.widths = list(axis.key.padding = 0,
                                                     ylab.right = 2)))

#plot histogram of median pirgd
hist(pirgd_midrange_bb[pirgd_midrange_bb >= 0 & pirgd_midrange_bb <= 250], main = "Midrange of PIRGd in SW Wyoming from 2002-2018", xlab = "DOY",
     xlim = c(0, 250), breaks = seq(0,250,10))

#remove variables
rm(pirgd_max, pirgd_min, pirgd_midrange, pirgd_midrange_bb)

```



