---
title: "Summary Statistics of PIRGd in WY"
author: "Ethan Berman"
date: "3/3/2020"
output: html_document
---

```{r, setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '/Volumes/SSD/climate_effects')

```

## Overview

This document contains summary statistics of PIRGd in WY from 2002-2018 from the eMODIS phenology dataset. Stats are calculated over the entire state and a smaller sub-region in SW Wyoming corresponding to sagebrush treatments.

## Mean PIRGd in Wyoming and SW Wyoming

```{r, echo=FALSE, message=FALSE}
#load packages
library(rgdal)
library(sp)
library(raster)

#load wyoming shapefile
wy <- readOGR('./reference/wyoming.shp', verbose = FALSE)

#transformto CRS of emodis
wy <- spTransform(wy, "+proj=laea +lat_0=45 +lon_0=-100 +x_0=0 +y_0=0 +a=6370997 +b=6370997 +units=m +no_defs")

#load PIRGd brick
pirgd <- brick('./emodis/pirgd/pirgd_wy_2002_2018.tif')

#mask to WY before calc summary stats
pirgd <- mask(pirgd, wy)

#calc mean pirgd for each pixel
pirgd_m <- calc(pirgd, function(x) {mean(x,na.rm = T)})

#calc number of na values in annual layers
pirgd_na <- calc(pirgd, function(x) {sum(is.na(x))})

#only save pirgd mean value if less than half of years are missing
#we have 17 years total
pirgd_m[pirgd_na > 8] <- NA

#round to whole DOY
pirgd_m <- round(pirgd_m)

#write pirgd mean raster
#writeRaster(pirgd_m, './emodis/pirgd/pirgd_mean_wy_2002_2018.tif', format = "GTiff", overwrite = T)

#load shapefile of treatments in SW WY
treat <- readOGR('/Users/Ediz/Team Braintree Dropbox/Ethan Berman/USGS/forage project/Phase 2/Initial Treatment Polygons/Project_Plannig_Areas_SWRMDCopy2.shp', verbose = FALSE)

#transform to CRS of emodis
treat <- spTransform(treat, "+proj=laea +lat_0=45 +lon_0=-100 +x_0=0 +y_0=0 +a=6370997 +b=6370997 +units=m +no_defs")

#plot treatments in WY
plot(wy, main = "Treatment Polygons in SW Wyoming")
plot(treat, add = T)

#calculate treatment bounding box
treat_bb <- bbox(treat)
treat_bb <- as(raster::extent(treat_bb), "SpatialPolygons")
proj4string(treat_bb) <- "+proj=laea +lat_0=45 +lon_0=-100 +x_0=0 +y_0=0 +a=6370997 +b=6370997 +units=m +no_defs"

#buffer at 2 km
treat_bb <- buffer(treat_bb, width = 2000)

#crop to WY
treat_bb <- crop(treat_bb, wy)

#plot mean PIRGd in WY
plot(wy, main = "Mean PIRGd in Wyoming from 2002-2018")
plot(pirgd_m, add = T,
     legend.args = list(text = 'DOY', side = 1, 
         font = 2, line = 0.5, cex = 1))
plot(treat_bb, add = T)

#plot histogram of mean pirgd
hist(pirgd_m, main = "Mean PIRGd in Wyoming from 2002-2018", xlab = "DOY")

#plot boxplot of mean pirgd
boxplot(pirgd_m, ylab = "DOY", main = "Boxplot of Mean PIRGd dates in WY from 2002-2018")

#crop mean pirgd to this bb
pirgd_m_bb <- crop(pirgd_m, treat_bb)

#plot pirgd mean raster for SW Wyoming
plot(pirgd_m_bb, main = "Mean PIRGd in SW Wyoming from 2002-2018",
     legend.args = list(text = 'DOY', side = 1, 
         font = 2, line = 0.5, cex = 1))

#plot histogram of mean pirgd
hist(pirgd_m_bb, main = "Mean PIRGd in SW Wyoming from 2002-2018", xlab = "DOY")

#remove variables
rm(pirgd_m, pirgd_m_bb)

```

## Median PIRGd in Wyoming and SW Wyoming

```{r, echo=FALSE, message=FALSE}
#calc median pirgd for each pixel
pirgd_med <- calc(pirgd, function(x) {median(x,na.rm = T)})

#only save pirgd mean value if less than half of years are missing
#we have 17 years total
pirgd_med[pirgd_na > 8] <- NA

#round to whole DOY
pirgd_med <- round(pirgd_med)

#write pirgd median raster
#writeRaster(pirgd_med, './emodis/pirgd/pirgd_median_wy_2002_2018.tif', format = "GTiff", overwrite = F)

#plot median PIRGd in WY
plot(wy, main = "Median PIRGd in Wyoming from 2002-2018")
plot(pirgd_med, add = T,
     legend.args = list(text = 'DOY', side = 1, 
         font = 2, line = 0.5, cex = 1))
plot(treat_bb, add = T)

#plot histogram of median pirgd
hist(pirgd_med, main = "Median PIRGd in Wyoming from 2002-2018", xlab = "DOY")

#plot boxplot of median pirgd
boxplot(pirgd_med, ylab = "DOY", main = "Boxplot of Median PIRGd dates in WY from 2002-2018")

#crop median pirgd to this bb
pirgd_med_bb <- crop(pirgd_med, treat_bb)

#plot pirgd mean raster for SW Wyoming
plot(pirgd_med_bb, main = "Median PIRGd in SW Wyoming from 2002-2018",
     legend.args = list(text = 'DOY', side = 1, 
         font = 2, line = 0.5, cex = 1))

#plot histogram of mean pirgd
hist(pirgd_med_bb, main = "Median PIRGd in SW Wyoming from 2002-2018", xlab = "DOY")

#remove variables
rm(pirgd_med, pirgd_med_bb)

```

## Variance of PIRGd in Wyoming and SW Wyoming

Variance > 5000 set to NA.

```{r, echo=FALSE, message=FALSE}
#calc var pirgd for each pixel
pirgd_var <- calc(pirgd, function(x) {var(x,na.rm = T)})

#only save pirgd var value if less than half of years are missing
#we have 17 years total
pirgd_var[pirgd_na > 8] <- NA

#round to whole DOY
pirgd_var <- round(pirgd_var)

pirgd_var2 <- pirgd_var
pirgd_var2[pirgd_var2 > 5000] <- NA
#write pirgd var raster
#writeRaster(pirgd_var, './emodis/pirgd/pirgd_var_wy_2002_2018.tif', format = "GTiff", overwrite = F)

#plot var PIRGd in WY
plot(wy, main = "Variance of PIRGd in Wyoming from 2002-2018")
plot(pirgd_var2, add = T,
     legend.args = list(text = 'Days', side = 1, 
         font = 2, line = 0.5, cex = 1))
plot(treat_bb, add = T)

#plot histogram of var pirgd
hist(pirgd_var2, main = "Variance of PIRGd in Wyoming from 2002-2018", xlab = "Days")

#crop var pirgd to this bb
pirgd_var_bb <- crop(pirgd_var2, treat_bb)

#plot pirgd var raster for SW Wyoming
plot(pirgd_var_bb, main = "Variance of PIRGd in SW Wyoming from 2002-2018",
     legend.args = list(text = 'Days', side = 1, 
         font = 2, line = 0.5, cex = 1))

#plot histogram of mean pirgd
hist(pirgd_var_bb, main = "Variance of PIRGd in SW Wyoming from 2002-2018", xlab = "DOY")

#remove variables
rm(pirgd_var, pirgd_var2, pirgd_var_bb)
```

## Mid-Range of PIRGd in Wyoming and SW Wyoming

Min + (Max-Min)/2

```{r, echo=FALSE, message=FALSE}
#calc mid pirgd for each pixel
pirgd_max <- calc(pirgd, function(x) {
        if(sum(is.na(x)) < 17) {max(x, na.rm = T)} else{NA}
        })

pirgd_min <- calc(pirgd, function(x) {
        if(sum(is.na(x)) < 17) {min(x, na.rm = T)} else{NA}
        })

pirgd_midrange <- pirgd_min + (pirgd_max - pirgd_min)/2

#only save pirgd var value if less than half of years are missing
#we have 17 years total
pirgd_midrange[pirgd_na > 8] <- NA

#round to whole DOY
pirgd_midrange <- round(pirgd_midrange)

#write pirgd midrange raster
#writeRaster(pirgd_midrange, '/Volumes/SSD/climate_effects/emodis/pirgd/pirgd_midrange_wy_2002_2018.tif', format = "GTiff", overwrite = F)

#plot var PIRGd in WY
plot(wy, main = "Midrange of PIRGd in Wyoming from 2002-2018")
plot(pirgd_midrange, add = T,
     legend.args = list(text = 'DOY', side = 1, 
         font = 2, line = 0.5, cex = 1))
plot(treat_bb, add = T)

#plot histogram of var pirgd
hist(pirgd_midrange, main = "Midrange of PIRGd in Wyoming from 2002-2018", xlab = "Days")

#crop var pirgd to this bb
pirgd_midrange_bb <- crop(pirgd_midrange, treat_bb)

#plot pirgd var raster for SW Wyoming
plot(pirgd_midrange_bb, main = "Midrange of PIRGd in SW Wyoming from 2002-2018",
     legend.args = list(text = 'DOY', side = 1, 
         font = 2, line = 0.5, cex = 1))

#plot histogram of mean pirgd
hist(pirgd_midrange_bb, main = "Midrange of PIRGd in SW Wyoming from 2002-2018", xlab = "DOY")

#remove variables
rm(pirgd_max, pirgd_min, pirgd_midrange, pirgd_midrange_bb)

```



